%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
\begin{document}
    \begin{abstract}
    \noindent \textbf{Index Terms:} Interactive Robot Learning , Reinforcement Learning with Feedback , learning from demonstration , Imitation Learning ,Human-Robot Interaction (HRI)\\
    \\
    \newline
    \noindent \textbf{Abstract}\\
    \\
    A new approach to developing skills for robots is interactive robot learning, which involves direct interaction with people and their environment.  In order to look into how robots can learn through interaction, this study analyzes 20 significant research papers published between 2017 and 2025.  The study explores a variety of learning strategies, including human provided feedback in reinforcement learning, imitation of human behavior, learning from models, and collaborative learning.  Important topics covered include teaching robots in natural language, integrating various communication methods, correcting errors instantly, and developing learning adaptive systems.  The study demonstrates significant advancements in both the methods by which humans can teach robots and the ways in which robots can learn from humans.  Manufacturing, social robots that engage with humans, and common consumer goods all benefit from these advancements.  This study highlights current problems such as scaling up these systems, ensuring their safety, and assisting robots in applying their acquired knowledge to new environments.  It also suggests interesting future directions for improving human robot collaborations that enhance robot capabilities and human machine collaboration.
    \end{abstract}
    


\end{document}
