%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
\begin{document}
    \chapter{Description of the Subject}
Rather than depending only on pre programmed behaviors or offline training datasets, the state of the art field of robotics known as Interactive Robot Learning (IRL) aims to enable robots to learn new technologies and information through direct human interaction.  Because IRL systems are made to continuously evolve and adapt under constant human guidance, they are much more valuable in dynamic environments like manufacturing, healthcare, assistive robotics, and social interaction scenarios than traditional robotic systems, which have fixed capabilities.  [11], [13], [15], [22], and [30].\\
 \\Establishing efficient channels of communication between humans and robots to enable significant knowledge transfer is a fundamental challenge in IRL.  Researchers have created a variety of interaction modalities to address this, such as brain computer interfaces, haptic feedback, gesture based interfaces, and natural language instructions.  These methods seek to establish simple and effective ways for people to teach robots new tasks while preserving the robot's capacity to apply previously learned behaviors in unfamiliar contexts.  [14], [16], [22], [31], [35], and [36].\\\\
 The annotated bibliography is divided into four main categories according to each paper's methodological focus in order to promote clarity and ease comparison.  One fundamental method is Learning from Demonstration, in which robots pick up skills by watching and imitating human behavior.  By concentrating on identifying the underlying patterns and principles from demonstrated examples in real world scenarios, these techniques allow robots to learn complex behaviors without explicit programming [3], [6], [10], [18], [20], and [25].\\\\
 Robots in the Interactive Reinforcement Learning category learn by making mistakes and getting constant feedback from human supervisors.  These methods speed up learning and guarantee appropriate, safe behaviors by fusing human guidance with the exploration powers of reinforcement learning.  Robots can navigate intricate reward systems and steer clear of potentially hazardous or suboptimal actions with the aid of human feedback [2], [8], [19], [23], [26], [30], and [36].\\\\
 The goal of human robot interaction architectures is to create all encompassing frameworks that facilitate ongoing learning interactions between people and robots.  These systems build strong platforms for interactive learning by combining adaptive interfaces, multimodal communication, and ontology based knowledge representation.  For learning processes to remain coherent over long periods of interaction, such architectures are necessary [4, 9, 13, 15, 17, 28, 29, 39].\\\\
 Domain specific implementations and innovative methodological advancements in interactive robot learning are included in the category of Specialized Applications and Techniques.  Predictive coding for body schema learning, physical interaction learning, dialog systems, and active learning techniques that enhance the human teaching process are all included in this.  [1], [5], [12], [21], [24], [27], [32], [33], [34], [37], [38], [40], [41].\\\\
 Beyond the techniques themselves, it's critical to emphasize how collaborative platforms and standardized assessment frameworks contribute to the advancement of interactive robot learning research.  By giving researchers reproducible experimental setups and common benchmarks, these tools facilitate systematic comparison of various methodologies and promote knowledge sharing among researchers.  Additionally, they aid in the creation of more resilient and broadly applicable systems that can function well in a variety of real world settings [7], [13].
\end{document}
